|

### 2023-07-20 paper/00-main.Rmd
revision submitted to arxiv

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
  # fig.pos = "t"
  # cache = TRUE
)
library(here)
library(tidyverse)
```

@@

\bigskip
\begin{center}
{\large\bf SUPPLEMENTARY MATERIAL}
\end{center}

>>>>

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
  # fig.pos = "p"
  # cache = TRUE
)
library(here)
library(tidyverse)
```

@@

<!-- \bigskip -->
\newpage
\begin{center}
{\large\bf SUPPLEMENTARY MATERIAL}
\end{center}

### 2023-07-20 paper/01-introduction.Rmd
revision submitted to arxiv

In an attempt to foster analytics and innovate the game, the National Football League (NFL) introduced their player tracking system known as Next Gen Stats (NGS) in 2016 [@nfl2023ngs].
NGS uses radio frequency identification (RFID) chips placed in players' shoulder pads (and in the ball) to collect data at a rate of 10 frames per second.
The data captures real-time on-field information such as locations, speeds, and accelerations of all 22 players (and the football).
While these data were initially only available for teams, media, and vendors, in December 2018 the NFL launched the inaugural edition of their annual Big Data Bowl (BDB) competition [@nfl2023big].

The first BDB led to several contributions largely focused on offensive performance evaluation. For example, one group of finalists introduced an approach for modeling the hypothetical completion probability of a pass aiding in the evaluation of quarterback (QB) decision making [@deshpande2020expected]. The winners of the inaugural BDB focused on identifying receiver routes via clustering techniques [@chu2020route] and convolutional neural networks [@sterken2019routenet]. Along with the BDB competition entries, the public release of NGS data allowed researchers to tackle <!-- (pun intended.  #eyeroll) --> a variety of other problems such as revisiting fourth down decision making [@lopez2020bigger], annotating pass coverage with Gaussian mixture models [@dutta2020unsupervised], and introducing a continuous-time framework to estimate within-play value [@yurko2020going].  Since its inception, the BDB has chosen a different theme each year leading to new insight about evaluating different positions such as running backs, defensive backs, and special teams. The 2023 edition of the NFL Big Data Bowl asked participants to evaluate linemen on passing plays [@Howard2023NFL].

Our focus of this manuscript is specifically on measuring the performance of defensive linemen in the NFL.
There are two main types of defensive linemen in American football: (1) defensive tackles (DTs) and (2) defensive ends (DEs).
Typically, DTs and DEs are located within the interior of the line and along the edges, respectively; see Figure \ref{fig:formation} (top) for example formation with DTs and DEs.
The primary purpose of both positions is to rush the QB on passing plays, with DEs displaying superiority in observed pass rushing ability [@Eager2018nfl].
Additionally, within DTs there are nose tackles (NTs) who directly line up across from the ball at the line of scrimmage; see Figure \ref{fig:formation} (bottom) for example defensive scheme with a NT.
NFL teams often employ either one NT or two DTs on the interior with DEs along either side of the defensive line.
In addition to defensive lineman, other positions may attempt to rush the QB on blitzing plays such as outside linebackers (OLBs), interior linebackers (ILBs), and potentially members of the secondary (cornerback, free safety, strong safety) whose primary role is pass coverage.
Note that apart from the formations shown in Figure \ref{fig:formation}, defensive linemen can have the flexibility to line up differently.
For example, a defensive end, depending on the opposing matchup, may not necessarily be positioned toward the outside of the line of scrimmage.

```{r formation, fig.width=4.8, fig.height=3.5, fig.pos="t", fig.cap="Two common defensive alignments in football: 4-3 defense (top) and 3-4 defense (bottom). The dashed line represents the line of scrimmage separating the defense and offense. Cornerbacks (CB) and safeties (S) are primarily involved in defending against passing plays.  Outside linbackers (OLB) and inside linebackers (ILB) generally play directly behind the defensive line and are involved in defending against passing and rushing plays.  Defensive tackles (DT), defensive ends (DE), and nose tackles (NT) primarily rush the quarterback on passing plays and attempt to stop the rusher as quickly as possible on running plays."}
tibble(x = c(3:5, 3:6 - 0.5, c(2, 3.7, 4.3, 6)),
       y = c(1, 1, 1, 1.01, 1.01, 1.01, 1.01, 1.02, 1.02, 1.02, 1.02),
       pos = c("DE", "NT", "DE", "OLB", "ILB", "ILB", "OLB",
               "CB", "S", "S", "CB"),
       d = "3-4 Defense") |> 
  bind_rows(
    tibble(x = c(3:6, c(2.5, 4.5, 6.5), c(2, 4.1, 4.9, 7)),
           y = c(1, 1, 1, 1, 1.01, 1.01, 1.01, 1.02, 1.02, 1.02, 1.02),
           pos = c("DE", "DT", "DT", "DE", "OLB", "MLB", "OLB",
                   "CB", "S", "S", "CB"),
           d = "4-3 Defense") 
  ) |> 
  mutate(d = fct_relevel(d, "4-3 Defense", "3-4 Defense")) |> 
  ggplot(aes(x, y)) +
  geom_tile(fill = "white") +
  geom_label(aes(label = pos), size = rel(4)) +
  geom_hline(yintercept = 0.995, linetype = "dashed") +
  facet_wrap(~ d, scales = "free", ncol = 1) +
  coord_cartesian(ylim = c(0.99, 1.03)) +
  theme_light() +
  theme(strip.text = element_text(size = rel(1.2)),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank())
```

In this work, using data made available in the Big Data Bowl 2023, we present a novel approach to measure the performance of pass rushers. Relative to other aspects of American football, such as QB evaluation [@burke2019deepqb; @reyers2021quarterback], the literature on evaluating pass rushers is scarce. Below, we provide a brief overview of existing pass rush metrics.

@@

More recently proposed metrics such as time in pocket (TIP), NGS get off [@Hermsmeyer2021nfl], and pass rush win rate (PRWR) [@Burke2018Created] are substantial improvements over the less sophisticated counting statistics, but nevertheless are still imperfect.
TIP refers to how long a QB can operate within the protected space behind the offensive line, known as the pocket.
However, this measure is highly context-dependent, as it can be influenced by a number of factors such as the defensive scheme or type of passing route.
NGS get off is an aggregated statistic, illustrating how quickly a defender can get past the line of scrimmage after the snap on average.
PRWR is created using player tracking data, which is at a more granular level than previous measures.
It demonstrates whether a pass rusher is able to beat their blocking matchup before a fixed time from the snap (2.5 seconds as chosen by ESPN).
However, this depends on the rather arbitrary time threshold used to define a pass rush win.
Besides, once a cutoff is chosen, PRWR converts continuous data to a win-loss indicator, becoming dichotomous like most of the metrics discussed above.

@@

The peer-reviewed literature on measuring the performance of football linemen (either offensive or defensive) is scant.
@AlamarGould2008 find an association between pass completion rate and successful pass blocking by offensive linemen.
The data for this study are collected for the first three weeks of the 2007 NFL season, manually recording whether a lineman holds a block and the time it took for quarterback to throw the football.
@AlamarGoldner2011 later follow up by using manually-tracked data for the 2010 season to estimate lineman performance for different team-positions instead of individual defenders (e.g., Chicago Bears' left tackle, Pittsburgh Steelers' center, etc.).
They used survival analysis to model time in pocket and completion percentage for quarterbacks before proposing a measure for linemen's contribution to their team's passing in terms of yards gained.

@@

In this paper, we focus on the evaluation of defensive linemen in football.
We propose STRAIN, a metric for measuring pass rush effectiveness, inspired by the concept of strain rate in materials science.
Our statistic gives a continuous measure of pressure for every pass rusher on the football field over the course of an entire play.
This allows for the assessment of pass rush success even on plays that do not result in an observed outcome like a sack, hit, or hurry.
We view this as a major step forward for accurately evaluating defensive linemen performance.
We also demonstrate that STRAIN is a stable quantity over time and predictive of defensive pressure.
Additionally, we consider a multilevel model to estimate every pass rusher's contribution to the average STRAIN in a play while controlling for team and various play-level information.
We note that although our focus in this paper is on pass rushers, our approach can be extended to the evaluation of pass blockers in American football.

>>>>

In an attempt to foster analytics and innovate the game, the National Football League (NFL) introduced their player tracking system known as Next Gen Stats in 2016 [@nfl2023ngs].
Next Gen Stats uses radio frequency identification (RFID) chips placed in players' shoulder pads (and in the ball) to collect data at a rate of 10 frames per second.
The data captures real-time on-field information such as locations, speeds, and accelerations of all 22 players (and the football).
While these data were initially only available for teams, media, and vendors, in December 2018 the NFL launched the inaugural edition of their annual Big Data Bowl competition [@nfl2023big].

The first Big Data Bowl led to several contributions largely focused on offensive performance evaluation. For example, one group of finalists introduced an approach for modeling the hypothetical completion probability of a pass aiding in the evaluation of quarterback (QB) decision making [@deshpande2020expected]. The winners of the inaugural Big Data Bowl focused on identifying receiver routes via clustering techniques [@chu2020route] and convolutional neural networks [@sterken2019routenet]. Along with the competition entries, the public release of NGS data allowed researchers to tackle <!-- (pun intended.  #eyeroll) --> a variety of other problems such as revisiting fourth down decision making [@lopez2020bigger], annotating pass coverage with Gaussian mixture models [@dutta2020unsupervised], and introducing a continuous-time framework to estimate within-play value [@yurko2020going].  Since its inception, the Big Data Bowl has chosen a different theme each year leading to new insight about evaluating different positions such as running backs, defensive backs, and special teams. The 2023 edition of the NFL Big Data Bowl asked participants to evaluate linemen on passing plays [@Howard2023NFL].

Our focus of this manuscript is specifically on measuring the performance of defensive linemen in the NFL.
There are two main types of defensive linemen in American football: defensive tackles and defensive ends.
Typically, these positions are located within the interior of the line and along the edges, respectively; see Figure \ref{fig:formation} (top) for example formation with defensive tackles and defensive ends.
The primary purpose of both positions is to rush the QB on passing plays, with defensive ends displaying superiority in observed pass rushing ability [@Eager2018nfl].
Additionally, within defensive tackles there are nose tackles who directly line up across from the ball at the line of scrimmage; see Figure \ref{fig:formation} (bottom) for example defensive scheme with a nose tackle.
NFL teams often employ either one nose tackle or two defensive tackles on the interior with defensive ends along either side of the defensive line.
Besides defensive lineman, other positions may attempt to rush the QB on blitzing plays such as outside linebackers, interior linebackers, and potentially members of the secondary (cornerbacks, free safeties, and strong safeties) whose primary role is pass coverage.
Note that apart from the formations shown in Figure \ref{fig:formation}, defensive linemen can have the flexibility to line up differently.
For example, a defensive end, depending on the opposing matchup, may not necessarily be positioned toward the outside of the line of scrimmage.

```{r formation, fig.width=4.8, fig.height=3.5, fig.pos="t", fig.cap="Two common defensive alignments in football: 4-3 defense (top) and 3-4 defense (bottom). The dashed line represents the line of scrimmage separating the defense and offense. Defensive tackles (DT), defensive ends (DE), and nose tackles (NT) primarily rush the QB on passing plays and attempt to stop the ball carrier as quickly as possible on running plays. Outside linebackers (OLB) and inside linebackers (ILB) usually play directly behind the defensive line and are involved in defending against passing and rushing plays. Cornerbacks (CB) and safeties (S) generally involve in defending against passing plays."}
tibble(x = c(3:5, 3:6 - 0.5, c(2, 3.7, 4.3, 6)),
       y = c(1, 1, 1, 1.01, 1.01, 1.01, 1.01, 1.02, 1.02, 1.02, 1.02),
       pos = c("DE", "NT", "DE", "OLB", "ILB", "ILB", "OLB",
               "CB", "S", "S", "CB"),
       d = "3-4 Defense") |> 
  bind_rows(
    tibble(x = c(3:6, c(2.5, 4.5, 6.5), c(2, 4.1, 4.9, 7)),
           y = c(1, 1, 1, 1, 1.01, 1.01, 1.01, 1.02, 1.02, 1.02, 1.02),
           pos = c("DE", "DT", "DT", "DE", "OLB", "MLB", "OLB",
                   "CB", "S", "S", "CB"),
           d = "4-3 Defense") 
  ) |> 
  mutate(d = fct_relevel(d, "4-3 Defense", "3-4 Defense")) |> 
  ggplot(aes(x, y)) +
  geom_tile(fill = "white") +
  geom_label(aes(label = pos), size = rel(4)) +
  geom_hline(yintercept = 0.995, linetype = "dashed") +
  facet_wrap(~ d, scales = "free", ncol = 1) +
  coord_cartesian(ylim = c(0.99, 1.03)) +
  theme_light() +
  theme(strip.text = element_text(size = rel(1.2)),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank())
```

In this work, using data made available in the Big Data Bowl 2023, we present a novel approach to measure the performance of pass rushers.
Relative to other aspects of American football, such as quarterback evaluation [@burke2019deepqb; @reyers2021quarterback], the literature on evaluating pass rushers is scarce.
Below, we provide a brief overview of existing pass rush metrics.

@@

More recently proposed metrics such as time in pocket, NGS get off [@Hermsmeyer2021nfl], and pass rush win rate [@Burke2018Created] are substantial improvements over the less sophisticated counting statistics, but nevertheless are still imperfect.
Time in pocket refers to how long a QB can operate within the protected space behind the offensive line, known as the pocket.
However, this measure is highly context-dependent, as it can be influenced by a number of factors such as the defensive scheme or type of passing route.
NGS get off is an aggregated statistic, illustrating how quickly a defender can get past the line of scrimmage after the snap on average.
Pass rush win rate is created using player tracking data, which is at a more granular level than previous measures.
It demonstrates whether a pass rusher is able to beat their blocking matchup before a fixed time from the snap (2.5 seconds as chosen by ESPN).
However, this depends on the rather arbitrary time threshold used to define a pass rush win.
Besides, once a cutoff is chosen, pass rush win rate converts continuous data to a win-loss indicator, becoming dichotomous like most of the metrics discussed above.

@@

The peer-reviewed literature on measuring the performance of football linemen (either offensive or defensive) is scant.
@AlamarGould2008 find an association between pass completion rate and successful pass blocking by offensive linemen.
The data for this study are collected for the first three weeks of the 2007 NFL season, manually recording whether a lineman holds a block and the time it took for quarterback to throw the football.
@AlamarGoldner2011 later follow up by using manually-tracked data for the 2010 season to estimate lineman performance for different team-positions instead of individual defenders (e.g., Chicago Bears' left tackle, Pittsburgh Steelers' center, etc.).
This work uses survival analysis to model time in pocket and completion percentage for quarterbacks before proposing a measure for linemen's contribution to their team's passing in terms of yards gained.

@@

In this paper, we focus on the evaluation of defensive linemen in football.
We propose STRAIN, a metric for measuring pass rush effectiveness, inspired by the concept of strain rate in materials science.
Our statistic gives a continuous measure of pressure for every pass rusher on the football field over the course of an entire play.
This allows for the assessment of pass rush success even on plays that do not result in an observed outcome like a sack, hit, or hurry.
We view this as a major step forward for accurately evaluating defensive linemen performance.
We also demonstrate that STRAIN is a stable quantity over time and predictive of defensive pressure.
Additionally, we consider a multilevel model to estimate every pass rusher's contribution to the average STRAIN in a play while controlling for player positions, team, and various play-level information.
We note that although our focus in this paper is on pass rushers, our approach can be extended to the evaluation of pass blockers in American football.

### 2023-07-20 paper/02-data.Rmd
revision submitted to arxiv

Along with the tracking information, the 2023 BDB includes scouting data provided by Pro Football Focus (PFF). 
This contains manually-collected player-level information, such as the player's role (e.g., whether they are a pass rusher and pass blocker) and credited events (e.g., player is credited with hitting the QB on the play).
In this manuscript, we use the PFF data to identify 36,362 unique pass rush attempts by players designated in the "pass rush" role across all plays.
For context, Figure \ref{fig:pass_rushers_count} displays the distribution of the number of observed pass rushers involved in a play, ranging from 1 to 8 with 4 pass rushers (i.e., 4-man rush formation) as the most common value.
We also leverage the PFF scouting data to count how many hits, hurries, and sacks each pass rusher is credited with across the span of observed data. 
Additionally, we use the PFF player roles to identify the blocking matchup for each pass rusher in order to adjust for opponent strength, as discussed in Section \ref{sec:methods}.

```{r}
#| label: pass_rushers_count
#| fig.cap: "Distribution of the number of pass rushers on passing plays."
#| fig.width: 4.5
#| fig.height: 3
source(here("scripts", "12_fig_pass_rushers_count.R"))
pass_rushers_count
```

>>>>

```{r}
#| label: pass_rushers_count
#| fig.cap: "Distribution of the number of pass rushers on passing plays."
#| fig.width: 3
#| fig.height: 2.5
source(here("scripts", "12_fig_pass_rushers_count.R"))
pass_rushers_count
```

Along with the tracking information, the Big Data Bowl 2023 includes scouting data provided by Pro Football Focus (PFF). 
This contains manually-collected player-level information, such as the player's role (e.g., whether they are a pass rusher and pass blocker) and credited events (e.g., player is credited with hitting the QB on the play).
In this manuscript, we use the PFF data to identify 36,362 unique pass rush attempts by players designated in the "pass rush" role across all plays.
For context, Figure \ref{fig:pass_rushers_count} displays the distribution of the number of observed pass rushers involved in a play, ranging from 1 to 8 with 4 pass rushers (i.e., 4-man rush formation) as the most common value.
We also leverage this scouting data to count how many hits, hurries, and sacks each pass rusher is credited with across the span of observed data. 
Additionally, we use the PFF player roles to identify the blocking matchup for each pass rusher in order to adjust for opponent strength, as discussed in Section \ref{sec:methods}.

### 2023-07-20 paper/03-methods.Rmd
revision submitted to arxiv

In materials science, strain [@callister2018materials] is the deformation of a material from stress, showing the change in a material's length relative to its original length.
Formally, let $L(t)$ be the distance between any given two points of interest within a material at time $t$, and $L_0$ be the initial distance between those two points.
The strain for a material at time $t$ is defined as 
\begin{equation*}
\varepsilon(t)= \frac{L(t) - L_{0}}{L_{0}}.
\end{equation*}
Notice that this measure is unitless due to being a ratio of two quantities having the same units.

Accordingly, the strain rate of a material measures the change in its deformation with respect to time.
Mathematically, the strain rate of a material can be expressed as the derivative of its strain.
That is,
\begin{equation*}
\varepsilon'(t) = \frac{d \varepsilon}{dt} = \frac{v(t)}{L_{0}},
\end{equation*}
where $v(t)$ is the velocity at which the two points of interest within the material are moving away from or towards each other.
Whereas strain has no units, the strain rate is measured in inverse of time, usually 1/seconds.

@@

Moreover, since we observe STRAIN at every tenth of a second within each play, we can then compute the average STRAIN across all frames played for every pass rusher.
Formally, the average STRAIN, denoted by $\overline{\text{STRAIN}}$, for pass rusher $j$ involved in $n_j$ total plays across $T_j$ total frames is
\begin{equation*}
\overline{\text{STRAIN}}_{j} = \frac{1}{T_j} \sum_{t=1}^{T_j} \widehat{\text{STRAIN}}_{j}(t).
\end{equation*}
This can be helpful for player evaluation, as we determine the most effective pass rushers based on their average STRAIN values in Section \ref{sec:strainbar}.
We also use average STRAIN to assess different statistical properties of our metric in Section \ref{sec:statprop}.

@@

To this end, we fit a multilevel model to evaluate pass rushers' impact on the average STRAIN observed in a play, while accounting for their team on defense, the opposing team on offense, and their assigned pass blocker.
We identify the pass blocker linked with the pass rusher of interest using the data provided by PFF.
Since there can be multiple blockers matching up with a rusher, for simplicity, we consider the nearest blocker positioned to the pass rusher at the start of the play.
We use random intercepts for the two player groups: pass rushers as $R$ and nearest pass blockers as $B$, as well as for the two team groups: defense $D$ and offense $O$.
We also account for attributes about pass rusher $j$ in play $i$ through the covariate vector $\mathbf{x_{ij}}$, and estimate their respective coefficients $\boldsymbol \beta$ as fixed-effects. 
Our model for the average STRAIN by pass rusher $j$ on play $i$ is as follows.
\begin{equation*}
\begin{aligned}
\overline{\text{STRAIN}}_{ij} &\sim N(R_{j[i]} + B_{b[ij]} + D_{d[i]} + O_{o[i]} + \mathbf{x_{ij}} \boldsymbol{\beta}, \sigma^2), \text{ for } i = 1, \dots, n \text{ plays} \\
R_{j} &\sim N(\mu_R, \sigma^2_R), \text{ for } j = 1, \cdots, \text{ \# of pass rushers}, \\
B_{b} &\sim N(\mu_B, \sigma^2_B), \text{ for } b = 1, \cdots, \text{ \# of pass blockers}, \\
D_{d} &\sim N(\mu_D, \sigma^2_D), \text{ for } d = 1, \cdots, \text{ \# of defensive teams}, \\
O_{o} &\sim N(\mu_O, \sigma^2_O), \text{ for } o = 1, \cdots, \text{ \# of offensive teams}.
\end{aligned}
\end{equation*}

@@

In order to provide a measure of uncertainty for our random effects, we use a bootstrapping strategy similar to the approach in @yurko2019nflwar.
Specifically, we resample team drives within games, which preserves the fact that team schedules are fixed but allows for random variation in player usage since this is dependent on team decision making.
By resampling plays within the same drive together, our sampling is generating realistic simulated data in comparison to sampling individual plays.
For each bootstrapped dataset, we fit the aforementioned multilevel model to obtain a distribution of estimates for the considered player and team effects.

>>>>

In materials science, strain [@callister2018materials] is the deformation of a material from stress, showing the change in a material's length relative to its original length.
Formally, let $L(t)$ be the distance between any given two points of interest within a material at time $t$, and $L_0$ be the initial distance between those two points.
The strain for a material at time $t$ is defined as 
\begin{equation*}
\varepsilon(t)= \frac{L(t) - L_{0}}{L_{0}}.
\end{equation*}
Notice that this measure is unitless due to being a ratio of two quantities having the same unit.

Accordingly, the strain rate of a material measures the change in its deformation with respect to time.
Mathematically, the strain rate of a material can be expressed as the derivative of its strain.
That is,
\begin{equation*}
\varepsilon'(t) = \frac{d \varepsilon}{dt} = \frac{v(t)}{L_{0}},
\end{equation*}
where $v(t)$ is the velocity at which the two points of interest within the material are moving away from or towards each other.
Whereas strain has no units, the strain rate is measured in inverse of time, usually inverse second.

@@

Moreover, since we observe STRAIN at every tenth of a second within each play, we can then compute the average STRAIN across all frames played for every pass rusher.
Formally, the average STRAIN, denoted by $\overline{\text{STRAIN}}$, for pass rusher $j$ involved in $n_j$ total plays across $\sum_{i\in Z_j} T_i$ total frames, where $Z_j$ is the set of all plays with pass rusher $j$'s involvement, is
\begin{equation*}
\overline{\text{STRAIN}}_{j} = \frac{1}{\sum_{i\in Z_j} T_i} \sum_{i\in Z_j} \sum_{t=1}^{T_i} \widehat{\text{STRAIN}}_{ij}(t).
\end{equation*}
This can be helpful for player evaluation, as we determine the most effective pass rushers based on their average STRAIN values in Section \ref{sec:strainbar}.
We also use average STRAIN to assess different statistical properties of our metric in Section \ref{sec:statprop}.

@@

To this end, we fit a multilevel model to evaluate pass rushers' impact on the average STRAIN observed in a play, while accounting for their team on defense, the opposing team on offense, and their assigned pass blocker.
We identify the pass blocker linked with the pass rusher of interest using the scouting data provided by PFF as mentioned in Section \ref{sec:data}.
Since there can be multiple blockers matching up with a rusher, for simplicity, we consider the nearest blocker positioned to the pass rusher at the start of the play.
We use random intercepts for the two player groups: pass rushers as $R$ and nearest pass blockers as $B$, as well as for the two team groups: defense $D$ and offense $O$.
We also account for attributes about pass rusher $j$ in play $i$ through the covariate vector $\mathbf{x_{ij}}$, and estimate their respective coefficients $\boldsymbol \beta$ as fixed-effects. 
Our model for the average STRAIN by pass rusher $j$ on play $i$ is as follows.
\begin{equation*}
\begin{aligned}
\overline{\text{STRAIN}}_{ij} &\sim N(R_{j[i]} + B_{b[ij]} + D_{d[i]} + O_{o[i]} + \mathbf{x_{ij}} \boldsymbol{\beta}, \sigma^2), \text{ for } i = 1, \dots, n \text{ plays} \\
R_{j} &\sim N(\mu_R, \sigma^2_R), \text{ for } j = 1, \cdots, \text{ \# of pass rushers}, \\
B_{b} &\sim N(\mu_B, \sigma^2_B), \text{ for } b = 1, \cdots, \text{ \# of pass blockers}, \\
D_{d} &\sim N(\mu_D, \sigma^2_D), \text{ for } d = 1, \cdots, \text{ \# of defensive teams}, \\
O_{o} &\sim N(\mu_O, \sigma^2_O), \text{ for } o = 1, \cdots, \text{ \# of offensive teams}.
\end{aligned}
\end{equation*}

@@

In order to provide a measure of uncertainty for our random effects, we use a bootstrapping strategy similar to the approach in @yurko2019nflwar.
Specifically, we resample team drives within games, which preserves the fact that team schedules are fixed but allows for random variation in player usage since this is dependent on team decision making.
By resampling plays within the same drive together, this allows us to generate realistic simulated data in comparison to sampling individual plays.
For each bootstrapped dataset, we fit the aforementioned multilevel model to obtain a distribution of estimates for the considered player and team effects.
